### 2021-07-07 rad-icfp-2021/correctness.tex
some progress correctness

TODO: I think need to explain here the (co)monoid structure induced by $D$

\begin{tabular}{l c l}
    $\seml$!J!$^T$!var!$_{T;i} \semr$ &=& TODO \\
    $\seml$!J!$^T$!op!$_{T;n} \semr$ &=&  \\
    $\seml$!J!$^T$!pair!$_{T;A\times B} \semr$ &=& \\
    $\seml$!J!$^T$!proj!$_{T1;T2;T3} \semr$ &=& \\
    $\seml$!J!$^T$!map2!$_{T;x,y.e}\semr$  &=& \\
    $\seml$!J!$^T$!reduce!$_{T;x,y.e;e}\semr$  &=& \\
    $\seml$!<e1, e2>!$\semr$  &=& !<!$\seml$!e1!$\semr$, $\seml$!e2!$\semr$!>! \\
    $\seml$!e1!$\icomp$!e2!$\semr$ &=&  \\
 \end{tabular}
TODO: Do I need any kind of proposition saying that this makes sense?
Might need to define on types as well to be clearer: It's an identity on types functor...

@@

\begin{definition}[Mutlicat from concat]
    A multicategory $\catC$ naturally defines a concategory with the same objects as $\catC$ and
    with morphisms $A\to [B_1,\ldots,B_n]$ being $n$ morphisms $A\to Bi$ of $\catC$. 
\end{definition}

@@

\begin{definition}[Concat from multicat]
    TODO
\end{definition}

@@

What remains to explain now is the reverse mode transformation between Source UNF and Target UNF.
We construct a functor $D:\ConcatS\to\ConcatT$. 
This functor will easily be shown to compute reverse-mode derivatives.
Because terms of Source UNF in $\ConcatS$, we observe the effect of $D$ on them
and show that it matches the syntactic $\Dsynrevsymbol_{\rho}$ 
from Section~\ref{sub:Simple reverse mode transformation}.

@@

\begin{proposition}[semantic of syntactic D matches D lax functor]
    TODO
\end{proposition}

@@

Recall that the intuition from Source UNF is that it consists of terms of Source that also return their context.
From there, the intuition for the Jacobian of a primitive in Target UNF is that it should be the Jacobian of
the corresponding term in Target. This is easily checked for scalar operations.

\begin{proposition}
    \begin{center}
\begin{tabular}{r c l}
    !$\seml\UNFSymbol^{-1}(J^T$map2$_{x,y.e})\semr$! &=& !$J^T\seml\UNFSymbol^{-1}($map2$_{x,y.e})\semr$!\\
    !$\seml\UNFSymbol^{-1}(J^T$reduce$_{x,y.e_1;e_2})\semr$! &=& !$J^T\seml\UNFSymbol^{-1}($reduce$_{x,y.e_1;e_2})\semr$!
\end{tabular}
\end{center}
\end{proposition}

@@

\begin{proof}
    TODO
\end{proof}

By inspecting what that composition of transformations does on terms of Source, 
we show that this indeed gives computes the same as the transformation 
$\directD{\rho}{\Gamma}{Y}$ from Section~\ref{sub:Macro for pure reverse mode transformation}.

@@

\begin{proof}
    TODO
\end{proof}

>>>>

TODO: I think need to explain here the (co)monoid structure induced by $\Dsynrevsymbol_{\rho}$

\begin{tabular}{l c l}
    $\seml$!J!$^T$!var!$_{T;i} \semr$ &=& TODO \\
    $\seml$!J!$^T$!op!$_{T;n} \semr$ &=& $id_T\concatcomp J^Top_n$ \\
    $\seml$!J!$^T$!pair!$_{T;A\times B} \semr$ &=& $id_T\concatcomp$!<$\pi_1,\pi_2$>! \\
    $\seml$!J!$^T$!proj!$_{T1;T2;T3} \semr$ &=& $id_{T_1}\concatcomp\widehat{0}_{T_2}\concatcomp id_{T_3}$ \\
    $\seml$!J!$^T$!map2!$_{T;x,y.e}\semr$  &=& $id_T\concatcomp map2_{T;x,y.e}$ \\
    $\seml$!J!$^T$!reduce!$_{T;x,y.e_1;e_2}\semr$  &=& $id_T\concatcomp reduce_{T;x,y.e_1;e_2}$ \\
    $\seml$!<e$_1$, e$_2$>!$\semr$  &=& !<!$\seml$!e1!$\semr$, $\seml$!e2!$\semr$!>! \\
    $\seml$!e$_1$!$\icomp$!e$_2$!$\semr$ &=& $\Lambda$($id_\Gamma\concatcomp\seml$!e2!$\semr$; $\Lambda^{-1}$($\seml$!e1!$\semr$)) \\
 \end{tabular}

@@

\begin{definition}[Mutlicat from concat]
    TODO
\end{definition}

@@

\begin{definition}[Concat from multicat]
    A multicategory $\catC$ naturally defines a concategory with the same objects as $\catC$ and
    with morphisms $A\to [B_1,\ldots,B_n]$ being $n$ morphisms $A\to Bi$ of $\catC$. 
\end{definition}

@@

What remains to explain now is the reverse mode transformation between Source UNF and Target UNF.
We construct a functor $\Dsynrevsymbol_{\rho}:\ConcatS\to\ConcatT$. 
This functor will easily be shown to compute reverse-mode derivatives.
Because terms of Source UNF in $\ConcatS$, we observe the effect of $\Dsynrevsymbol_{\rho}$ on them
and show that it matches the syntactic $\Dsynrevsymbol_{\rho}$ 
from Section~\ref{sub:Simple reverse mode transformation}.

@@

The design of $\Dsynrevsymbol_{\rho}$ on UNF is inspired by the semantic one between concategories, 
and it's routine to check that they match.

\begin{proposition}[semantic of syntactic D matches D lax functor]
    Given a well typed term !T$_1$ $\vdash$ e: T$_2$! in Source UNF, we have
    \begin{center}
        $\seml\Dsynrevsymbol_{\rho}$!e!$\semr$ = $\Dsynrevsymbol_{\rho}\seml$!e!$\semr$
    \end{center}
\end{proposition}

@@

Recall that the intuition from Source UNF is that it consists of terms of Source that also return their context.
From there, the intuition for the Jacobian of a primitive in Target UNF is that it should be the Jacobian of
the corresponding term in Target. This is easily checked for scalar operations. For the more difficult cases, we have

\begin{proposition}
$\UNFSymbol^{-1}$ preserves the semantics of Jacobians of !map2! and !reduce!.
    \begin{center}
\begin{tabular}{r c l}
    !$\seml\UNFSymbol^{-1}(J^T$map2$_{x,y.e})\semr$! &=& !$J^T\seml\UNFSymbol^{-1}($map2$_{x,y.e})\semr$!\\
    !$\seml\UNFSymbol^{-1}(J^T$reduce$_{x,y.e_1;e_2})\semr$! &=& !$J^T\seml\UNFSymbol^{-1}($reduce$_{x,y.e_1;e_2})\semr$!
\end{tabular}
\end{center}
\end{proposition}

@@

By inspecting what that composition of transformations does on the terms of Source, 
we show that this indeed gives computes the same as the transformation 
$\directD{\rho}{\Gamma}{Y}$ from Section~\ref{sub:Macro for pure reverse mode transformation}. 
This should not come as a surprise because the design of $\directD{\rho}{\Gamma}{Y}$ was in fact guided via 
this decomposition and intermediate representation.

@@
